---
title: "Data Manipulation"
author: "Diana Ballesteros"
date: "9/20/2018" 
output: html_document
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```


##
```{r}

litters_data = read_csv("./data/FAS_litters.csv",
  col_types = "ccddiiii")
litters_data = janitor::clean_names(litters_data)

pups_data = read_csv("./data/FAS_pups.csv",
  col_types = "ciiiii")
pups_data = janitor::clean_names(pups_data)


```

##Select Variables 

```{r}
select(litters_data, group, litter_number, gd0_weight, pups_survive)
```

....by specifying a range: 
```{r}
select(litters_data, gd_df_birth:pups_survive)
select(litters_data, group, litter_number, gd0_weight, pups_survive)
```


...by saying you want to remove (this will allow you to remove a column you don't want): 
```{r}
select(litters_data, -litter_number)
```

...Selecting gd_day_0_weight and also renaming it as well. You can do either. 
```{r}
select(litters_data, group, litter_number, gd_day_0_weight = gd0_weight)
rename(litters_data,  gd_day_0_weight = gd0_weight)
```

....use select helpers

```{r}
?select_helpers

select(litters_data, starts_with("gd"))
select(litters_data, litter_numer, starts_with("pup"))

select(litters_data, litter_number, everything())
```

...select the columns containing litter number, sex, and PD ears.
```{r}
select(pups_data, litter_number:pd_ears)
```

##Filter observations

...gd_of_birth is exactly equal to 20 
```{r}
gd_of_birth == 20 
```

pups_born_alive >= 2
pups_survive != 4
!(pups_survive == 4)
group %in% c("Con7", "Con8")
group == "Con7" & gd_of_birth == 20
!is.na(wt_increase)

```{r}
select(litters_data, group, litter_number, gd_day_0_weight = gd0_weight)

filter(litters_data, gd0_weight < 25)
filter(litters_data, gd0_weight >= 25) 

filter(litters_data, pups_born_alive == 8) 
filter(litters_data, !is.na(gd0_weight)) 

filter(litters_data, group == "Low8" | group == "Low7")
filter(litters_data, group %in% c("Low8","Low7")) 
filter(litters_data, pups_data)

```

##Learning Assessment 
```{r}
select(pups_data, litter_number:pd_walk)
filter(pups_data, sex == 1)
filter(pups_data, pd_walk < 11, sex == 2)
```


##Mutate: you can multiple things at once, tolower allows you to do lower case you aren't necessary changing the table (only in the console) 
```{r}
mutate(litters_data,
  wt_gain = gd18_weight - gd0_weight,
  group = tolower(group)
)

mutate(litters_data,
  wt_gain = gd18_weight - gd0_weight,
  wt_gain_squared = wt_gain^2
)
```


##Arrange 

Arange the data...
Desc makes it in descending order
```{r}

arrange(litters_data, gd0_weight)

arrange(litters_data, gd0_weight, pups_born_alive)
arrange(litters_data, desc(gd0_weight), pups_born_alive)

head(arrange(litters_data, group, pups_born_alive), 10)
```

##Piping 

Look at intermediate

```{r}
litters_data_raw = read_csv("./data/FAS_litters.csv",
  col_types = "ccddiiii")
litters_data_clean_names = janitor::clean_names(litters_data_raw)
litters_data_selected_cols = select(litters_data_clean_names, -pups_survive)
litters_data_with_vars = mutate(litters_data_selected_cols, 
  wt_gain = gd18_weight - gd0_weight,
  group = tolower(group))
litters_data_with_vars

```

use nested funstion calls: it is being saved as litter_data_clean 
```{r}
litters_data_clean = 
  mutate(
    select(
      janitor::clean_names(
        read_csv("./data/FAS_litters.csv", col_types = "ccddiiii")
        ), 
    -pups_survive
    ),
  wt_gain = gd18_weight - gd0_weight,
  group = tolower(group)
  )
litters_data_clean
```

Use piping 
```{r}
litters_data = 
  read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names(dat = .) %>%
  select(.data = ., -pups_survive) %>%
  mutate(.data = .,
    wt_gain = gd18_weight - gd0_weight,
    group = tolower(group)) %>% 
  filter(!is.na(gd0_weight)) %>% 
  filter(.data = ., !is.na(gd_weight))
```

Illustrate placeholder use with "lm"
The dot is just a placeholder for the code prior 
```{r}
  read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names(dat = .) %>%
  select(.data = ., -pups_survive) %>%
  mutate(.data = .,
    wt_gain = gd18_weight - gd0_weight,
    group = tolower(group)) %>% 
  filter(!is.na(gd0_weight)) %>% 
  lm(gd18_weight ~ gd0_weight, data = .)
```


